Data Specialization for a Staged Lambda Calculus

### 2015-02-26 paper/splitting.tex
rewriting splitting rules explanation

When $e$ is unit or a variable,
splitting yields a combined term formed by tupling $e$ with a $\tup{}$ precomputation.
These expressions, as they contain no second-stage subcomputations, have a trivial resumer, $\tup{}$.
Similarly, the contents of a \texttt{grnd} block are assigned entirely to the immediate value,
with trivial precomputation and resumer.

For all non-terminals (except \texttt{next}),
splitting descends into $e$, recursively splitting its $n$ subterms
to produce their respective combined terms $c_1,\ldots,c_n$ and resumers $r_1, \ldots, r_n$.
The combined term of $e$ is formed by binding $c_1,\ldots,c_n$
to the patterns $(y_1,z_1),\ldots,(y_n,z_n)$
to isolate immediate results from boundary values. Then,
the immediate result of $e$ is formed by replacing $e$'s subterms with $y_1,\ldots,y_n$.
The resumer binds the boundary values $b_1,\ldots,b_n$ to an
argument $(l_1,\ldots,l_n)$ in a term that has the same structure
of~$e$ but where each subterm is replaced by its resumer ($r_i$'s).

>>>>

We start with splitting the unit value, $\tup{}$.
Masking tells us that units, being trivial, only split into more units: $\tup{} \vsplito \mval {\tup{}} {\tup{}}$.
Plugging this in to the contract for splitting at \bbonem\ gives:
\[
	\text{if } \splitone {\tup {}} A c {l.r} \text{ then } \reduce c {\tup{\tup{},b}} \text{ and } [b/l]r \equiv \tup{}
\]
Although the only hard requirement for $b$, $c$, and $l.r$ is to satisfy these relations,
our actual choices are designed to do the least work possible.
In particular, we choose $b = \tup{}$; $c=\tup{\tup{},\tup{}}$; and $l.r = \_.\tup{}$.
That is, we choose a trivial boundary value.
All together, this yield the rule in \cref{fig:termSplit}.
This exercise proceeds much the same way for \texttt{grnd} blocks, 
with the purely first stage contents of the block standing in for one of the $\tup{}$s.

We next consider splitting a non-terminal, namely $\inl {e}$.
From the semantics, we already know that 
for some $\xi$ and $v$, $\diaone {e} \xi v$ and $\diaone {\inl e} \xi {\inl v}$.
Masking these outputs gives $\rtab \xi v \vsplito \mval i q$ 
and $\rtab \xi {\inl v} \vsplito \mval {\inl i} q$.
Since splitting works from the inside out, we already know that
\[
	\splitone e A c {l.r} \text{ and } \reduce c {\tup{i,b}} \text{ and } [b/l]r \equiv q
\]
and what we need to do is find some $b'$, $c'$, and $l'.r'$ satisfying
\[
	\reduce {c'} {\tup{\inl i,b'}} \text{ and } [b'/l']r' \equiv q
\]
Once again, we take the simplest solution, which in this case is $b' = b$; $l'.r' = l.r$;
and $c' = \letin {\tup {x,y}} c {\tup{\inl x, y}}$.  
This sort of technique generalize to all constructs which don't involve control flow.

\nr{working from here down}
